Structure Learning for $\VAR(p)$ Models

### 2019-05-13 paper/paper.tex
Significant work on completing a converse of the main ancestor_properties result.  Introduced notion of a persistent system.

@@

\begin{proposition}[Invariance Under Invertible and Deterministic Modifications]
  Let $x(t)$ be a WSS process with absolutely summable covariance
  sequence, and spectral density uniformly bounded above and below.  Let
  $F(z), G$ represent univariate and invertible
  linear-time-invariant filters.  And, let $f(t), g(t)$ be perfectly
  predictable processes.  Then,

  \begin{equation}
    x_j \gc x_i \iff F(x_j - f)(t) \gc G(x_i - g)(t)
  \end{equation}

\begin{proof}
  \hl{TODO}
\end{proof}
\end{proposition}

@@

@@

\begin{definition}[Grandparents, ancestors]
The set of level $\ell$ \textit{grandparents} of node $i$, denoted
$\gpn{\ell}{i}$, is the set such that $j \in \gpn{\ell}{i}$ if and only if
there is a \textit{directed and cycle-free path} of length $\ell$ in
$\gcg$ from $j$ to $i$ \textit{or} if $j = i$ and there is a cycle of
length $\ell$ starting at $i$ and ending back at $i$, that is, $i$ can be
it's own grandparent.  Clearly, $\pa{i} = \gpn{1}{i}$.

Finally, the set of \textit{level $\ell$ ancestors} of
$i$: $\ancn{\ell}{i} = \bigcup_{\lambda \le \ell}\gpn{\lambda}{i}$ is the set such that
$j \in \ancn{\ell}{i}$ if and only if there is a directed and cycle-free
path of length $\ell$ \textit{or less} in $\gcg$ from $j$ to $i$
\textit{or} if $j = i$ and there is a directed cycle of length $\ell$ or
less from $i$ back to $i$ in $\gcg$.  The set of \textit{all
  ancestors} of $i$ (i.e. $\ancn{n}{i}$) is denoted simply $\anc{i}$.
\end{definition}

Even though the definitions of Granger-causality naturally suggests
that each node in the graph should have a self-loop (unless the node
is simply independent noise) we will not include nodes in their own
parent sets: $i \not\in \pa{i} \subseteq \anc{i}$.  However, we do not
immediately assume that $\gcg$ is a DAG, and so \textit{we will allow for $i$
to be it's own ancestor}.

@@

  \begin{equation}
    x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t).
  \end{equation}

@@

  where $\A(z) = \sum_{\tau = 0}^\infty A(\tau)$ is the filter from
  the Wold decomposition representation of $x(t)$, equation
  (\ref{eqn:wold}).
\end{proposition}

@@

  \begin{align*}
    x_i(t) &= (I - \B(z))^{-1}\\
    &= \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
    &= \A_{ii}(z) v_i(t) + \sum_{\substack{j \in \anc{i} \\ j \ne i}} \A_{ij}(z) v_j(t)\\
  \end{align*}
\end{proof}

@@

  where $\H_2 \vee \H_2 = \cl \{\alpha + \beta\ |\ \alpha \in \H_1, \beta \in \H_2 \}$ is the closed sum of subspaces.
\end{lemma}
\begin{proof}
  We can apply the Gram-schmidt process to the collection of subspaces to obtain a sequence of orthogonal subspaces $\widetilde{\H}_0, \widetilde{\H}_1, \ldots$ with $\widetilde{\H}_0 = \H_0$, $\widetilde{\H}_n \subseteq \H_n$, and $\bigvee_{n = 0}^\infty\H_n = \bigoplus_{n = 0}^\infty \widetilde{\H}_n$, where $\oplus$ indicates the orthogonal sum of subspaces (which is always closed).  Then, we have an explicit representation of the projection

@@

\begin{definition}[\hl{Confounder}]
  A node $k$ will be referred to as a \textit{confounder} of nodes
  $i, j$ (neither of which are equal to $k$) if
  $k \in \anc{i} \cap \anc{j}$ and there exists a path
  $\gcgpath{k}{i}$ not containing $j$, and a path $\gcgpath{k}{j}$
  not containing $i$.
\end{definition}

\begin{proposition}
  \label{prop:ancestor_properties}
  In a Granger-causality graph $\gcg$, if $j \pwgc i$ then at least
  one of the following must hold

@@

  The first part is used in part to prove the second (and stronger) statement, which allows common ancestors, but only if they are ``mediated'' by $i$.  For example, this covers the case of a graph $k \rightarrow i \rightarrow j$, where $\{k\} = \anc{i} \cap \anc{j}$, but clearly $j \npwgc i$.

  An ancestor such as that in the second part of the theorem can be referred to as a \textit{confounder}, and a simple example is given by the fork $i \leftarrow k \rightarrow j$.
\end{remark}

@@

\subsection{Strongly Causal Graphs}
The examples at the conclusion of the previous section seem rather
pathological since we have simply constructed a case where different
paths in the graph cancel exactly.  It may be possible to rid
ourselves of these pathologies for instance by defining a notion of
``robust'' pairwise causality where ``$j \pwgc _R i$'' if $j \pwgc i$
almost surely for a ``small'' random perturbation of the system
matrix.  However, we do not believe such a modification of our theory
can have useful practical implications since the difference between
paths that ``nearly cancel'' rather than cancel exactly would in
practice require inordinate amounts of data to resolve.

Alternatively, we could restrict the coefficients of the system
matrix, e.g. by requiring that $B_{ij}(\tau) \ge 0$.  Instead, we
think it is more meaningful to focus on the defining feature of time
series networks, that is, the topology of $\gcg$.

\begin{definition}[Strongly Causal]
  We will say that a Granger-causality graph $\gcg$ is \textit{strongly causal}
if there is at most 1 directed path between any two nodes.

\hl{TODO: Should DAG be included in this definition?  I think it is
  not necessary for some theorems.}
\end{definition}

Examples of strongly causal graphs include directed trees (or
forests), DAGs where each node has at most one parent, and each of the
figures of this paper.  Fortunately, the strongly causal property is
inherited by subgraphs:

@@

@@

@@

  \begin{align*}
    x_i(t) &= v_i(t) + \B_{ii}(z) x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z) x_k(t)\\
           &= \underbrace{v_{a_0}(t) + \B_{a_0a_0}(z) x_i(t) + \sum_{\substack{k \in \pa{a_0} \\ k \ne a_1}}\B_{a_0 k}(z) x_k(t)}_{\defeq \wtalpha{a_0}{a_1}} + \B_{a_0a_1}(z)x_{a_1}(t)\\
           &= \wtalpha{a_0}{a_1} + \B_{a_0a_1}(z)\big[\wtalpha{a_1}{a_2} + \B_{a_1a_2}(z)x_{a_2}(t) \big]\\
           &\overset{(a)}{=} \cdots\\
           &= \sum_{\ell = 0}^r \underbrace{\Big(\prod_{m = 0}^{\ell - 1} \B_{a_m a_{m + 1}}(z) \Big)}_{\defeq F_\ell(z)} \wtalpha{a_\ell}{a_{\ell + 1}} + \Big(\prod_{m = 0}^{r}\B_{a_m a_{m + 1}}(z)\Big)x_{a_{r + 1}}(t)\\
           &= \sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}} + F_{r + 1}(z) x_j(t)
  \end{align*}

@@

  but this is impossible since with probability one $x_i(t) \not \in \H_{t - 1}^{(i)}$.
\end{proof}

@@

@@

    \BlankLine
    \caption{Pairwise Graph Recovery}
    \label{alg:pwgr}
    % \TitleOfAlgo{Pairwise Graph Recovery}
    \Input{Pairwise Granger-causality relations between $n$ processes whose joint Granger-causality relations are known to form a strongly causal DAG $\gcg$.}
    \Output{Edges $E = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of the graph $\gcg$.}
    \Initialize{$S_0 = [n]$  \texttt{\# unprocessed nodes}\\
      $E = \emptyset$  \texttt{\# edges of }$\gcg$\\
      % $P_0 = \emptyset$  \texttt{\# layer by layer driving nodes}\\
      $k = 0$ \texttt{\# a counter used only for notation}}
    \BlankLine
    $W \leftarrow \{(i, j)\ |\ i \pwgc j, j \npwgc i\}$  \texttt{\# potential edges}\\
    \While{$S_k \ne \emptyset$}{
      $P_k \leftarrow \{i \in S_k\ |\ \forall s \in S_k\ (s, i) \not\in W\}$  \texttt{\# parents of current layer}\\
      % $P_k \leftarrow P_{k - 1} \cup P_k$ \texttt{\# Expand parent set}\\
      $S_{k + 1} \leftarrow S_k \setminus P_k$ \texttt{\# remove from } $S$\\

@@

      \For{$r = 0, \ldots, k$}
      {
        \label{alg:inner_loop}
        $D_{kr} \leftarrow \{(i, j) \in P_{k - r} \times C_k\ |\ (i, j) \in W,\ \text{no } i \rightarrow j \text{ path in } E \}$\\%  \texttt{\# edges incident on }$C_k$\\
        $E \leftarrow E \cup D_{kr}$ \texttt{\# add edges to }$E$\\
        % $W_{k + 1} \leftarrow W_k \setminus D_{kr}$  \texttt{\# remove edges from consideration}\\
      }
      % $W_{k + 1} \leftarrow W_k \setminus \big(\bigcup_{r = 0}^k D_{kr}\big)$  \texttt{\# remove edges from consideration}\\

      \;
      $k \leftarrow k + 1$
    }
    \Return{$E$}
  \end{algorithm}
\end{theorem}

@@

  The algorithm proceeds first with the parentless nodes $1, 2$ on the
  initial iteration where the edge $(1, 3)$ is added to $E$.  On the
  next iteration, the edges $(3, 4), (2, 4)$ are added, and the false
  edge $(1, 4)$ is excluded due to the path
  $1 \rightarrow 3 \rightarrow 4$ in E.  Finally, edges
  $(3, 5), (4, 6)$ are added, and the false
  $(1, 5), (1, 6), (3, 6), (2, 6)$ edges are similarly excluded due to
  the ordering of the inner loop.

  \begin{figure}
    \centering
    \caption{Example graph for Algorithm \ref{alg:pwgr}}
    \footnotesize{Black arrows indicate true parent-child
      relations.  Red dotted arrows indicate pairwise causality (due to
      ancestor relations), green dash-dotted arrow indicates
      bidirectional pairwise causality (due to the confounding node
      $1$).  Blue groupings indicate each $P_k$ in Algorithm
      \ref{alg:pwgr}.  Repetition is for aesthetics.}
    \label{fig:example_fig3}

@@

>>>>

% TODO: read this https://lemire.me/blog/rules-to-write-a-good-research-paper/

% TODO:
% -- define "DAG"
% -- remove speculative statements
% --

@@

% \begin{proposition}[Invariance Under Invertible and Deterministic Modifications]
%   Let $x(t)$ be a WSS process with absolutely summable covariance
%   sequence, and spectral density uniformly bounded above and below.  Let
%   $F(z), G$ represent univariate and invertible
%   linear-time-invariant filters.  And, let $f(t), g(t)$ be perfectly
%   predictable processes.  Then,

%   \begin{equation}
%     x_j \gc x_i \iff F(x_j - f)(t) \gc G(x_i - g)(t)
%   \end{equation}

% \begin{proof}
%   \hl{TODO}
% \end{proof}
% \end{proposition}

@@

\begin{definition}[Graph Theory Review]
  A \textit{graph} $\gcg = (\mathcal{V}, \mathcal{E})$ is simply a
  tuple of sets respectively called \textit{nodes} and \textit{edges}.
  Throughout this paper, we have in all cases
  $V = [n] \defeq \{1, 2, \ldots, n\}$.  We will also focus solely on
  \textit{directed} graphs, where the edges
  $\mathcal{E} \subseteq V \times V$ are \textit{ordered} pairs.

  A (directed) \textit{path} (of length $r$) between nodes $i, j$,
  denoted $\gcgpath{i}{j}$, is a sequence
  $a_0, a_1, \ldots, a_{r - 1}, a_r$ with $a_0 = i$ and $a_r = j$ such
  that $\forall\ 0 \le k \le r\ (a_k, a_{k + 1}) \in \mathcal{E}$.

  A \textit{cycle} is a path of length $2$ or more between a node and
  itself.  An edge between a node and itself $(i, i)$ (which is not a
  cycle) is referred to as a \textit{loop}.

  A graph $\gcg$ is a \textit{directed acyclic graph} (DAG) if it is a
  directed graph and does not contain any cycles.
\end{definition}

@@

\begin{definition}[Grandparents, ancestors]
  The set of level $\ell$ \textit{grandparents} of node $i$, denoted
  $\gpn{\ell}{i}$, is the set such that $j \in \gpn{\ell}{i}$ if and
  only if there is a \textit{directed path} of length $\ell$ in $\gcg$
  from $j$ to $i$.  Note that a node can be it's own grandparent
  unless $\gcg$ is a DAG.  Clearly, $\pa{i} = \gpn{1}{i}$.

  Finally, the set of \textit{level $\ell$ ancestors} of $i$:
  $\ancn{\ell}{i} = \bigcup_{\lambda \le \ell}\gpn{\lambda}{i}$ is the
  set such that $j \in \ancn{\ell}{i}$ if and only if there is a
  directed path of length $\ell$ \textit{or less} in $\gcg$ from $j$
  to $i$.  The set of \textit{all ancestors} of $i$
  (i.e. $\ancn{n}{i}$) is denoted simply $\anc{i}$.

  Note that since loops are not considered paths, it is not possible
  for a node to be it's own parent.
\end{definition}

@@

  \begin{equation}
    x_i(t) = \wtB_{ii}(z)v_i(t) + \sum_{k \in \pa{i}}\wtB_{ik}(z)x_k(t),
  \end{equation}

  where $\wtB_{ii}(z) = (1 - \B_{ii}(z))^{-1}$ and
  $\wtB_{ik}(z) = \wtB_{ii}(z) B_{ik}(z)$.

@@

  where $\A(z) = \sum_{\tau = 0}^\infty A(\tau)z^{-\tau}$ is the filter from
  the Wold decomposition representation of $x(t)$, equation
  (\ref{eqn:wold}).
\end{proposition}

@@

  \begin{align*}
    x_i(t) &= (I - \B(z))^{-1}v(t)\\
    &= \sum_{j = 1}^n \A_{ij}(z) v_j(t)\\
    &= \A_{ii}(z) v_i(t) + \sum_{\substack{j \in \anc{i} \\ j \ne i}} \A_{ij}(z) v_j(t)
  \end{align*}
\end{proof}

@@

  where $\H_1 \vee \H_2 = \cl \{\alpha + \beta\ |\ \alpha \in \H_1, \beta \in \H_2 \}$ is the closed sum of subspaces.
\end{lemma}
\begin{proof}
  We can apply the Gram-schmidt process to the collection of subspaces to obtain a sequence of orthogonal subspaces $\widetilde{\H}_0, \widetilde{\H}_1, \ldots$ with $\widetilde{\H}_0 = \H_0$, $\widetilde{\H}_n \subseteq \H_n$, and $\bigvee_{n = 0}^\infty\H_n = \bigoplus_{n = 0}^\infty \widetilde{\H}_n$, where $\oplus$ indicates the orthogonal sum of subspaces (which is always closed).  Then, we have an explicit representation of the projection

@@

\begin{definition}[\hl{Confounder}]
  A node $k$ will be referred to as a \textit{confounder} of nodes
  $i, j$ (neither of which are equal to $k$) if
  $k \in \anc{i} \cap \anc{j}$ and there exists a path
  $\gcgpath{k}{i}$ not containing $j$, and a path $\gcgpath{k}{j}$
  not containing $i$.

  A simple example is furnished by the ``fork'' graph
  $i \leftarrow k \rightarrow j$.
\end{definition}

\begin{proposition}
  \label{prop:ancestor_properties}
  \hl{Add the stronger statements on uncorrelatedness for use on confounders later.}

  In a Granger-causality graph $\gcg$, if $j \pwgc i$ then at least
  one of the following must hold

@@

  The first part is used in part to prove the second (and stronger) statement, which allows common ancestors, but only if they are ``mediated'' by $i$.  For example, this covers the case of a graph $k \rightarrow i \rightarrow j$, where $\{k\} = \anc{i} \cap \anc{j}$, but clearly $j \npwgc i$.
\end{remark}

@@

\subsection{Strongly Causal Graphs}
\label{sec:strongly_causal_graphs}
% The examples at the conclusion of the previous section seem rather
% pathological since we have simply constructed a case where different
% paths in the graph cancel exactly.  It may be possible to rid
% ourselves of these pathologies for instance by defining a notion of
% ``robust'' pairwise causality where ``$j \pwgc _R i$'' if $j \pwgc i$
% almost surely for a ``small'' random perturbation of the system
% matrix.  However, we do not believe such a modification of our theory
% can have useful practical implications since the difference between
% paths that ``nearly cancel'' rather than cancel exactly would in
% practice require inordinate amounts of data to resolve.

In this section and the next we will seek to understand when converse
statements of Proposition \ref{prop:ancestor_properties} \textit{do}
hold.  One possibility is to restrict to coefficients of the system
matrix matrix, e.g. by requiring that $B_{ij}(\tau) \ge 0$.  Instead,
we think it is more meaningful to focus on the defining feature of
time series networks, that is, the topology of $\gcg$.

\begin{definition}[Strongly Causal]
  \label{def:strongly_causal}
  We will say that a Granger-causality graph $\gcg$ is
  \textit{strongly causal} if there is at most 1 directed path between
  any two nodes.
\end{definition}

Examples of strongly causal graphs include directed trees (or
forests), DAGs where each node has at most one parent, and figure
\ref{fig:example_fig3} of this paper.  The persistence condition is
used to ensure that $\linE{x_i(t)}{\H_{t - 1}^{(i)}} \ne 0$ i.e. that
the each process has at least some amount of ``memory'' -- this can
occur either due to cycles, or non-zero $\B_{ii}(z)$ terms in equation
\eqref{eqn:ar_representation}.  It is evident that the strong causal property
is inherited by subgraphs, but persistence need not be, unless $\gcg$ is a DAG.

\hl{Not sure these are needed, but were important in my own thinking.}

@@

\begin{lemma}
  \label{lem:still_persistent}
  If $\gcg$ is a persistent DAG, any subgraph formed by eliminating nodes
  as well as all of the in and out edges thereof is still persistent.
\end{lemma}
\begin{proof}
  \hl{TODO -- is this needed?}
\end{proof}

@@

\begin{proposition}
  \label{prop:independent_confounders}
  If $\gcg$ is strongly causal we have distinct
  $k_1, k_2 \in \anc{i} \cap \anc{j}$ are confounders of $i, j$, then
  $k_1 \not\in \anc{k_2}$, $k_2 \not\in \anc{k_1}$ and
  $\not\exists \ell \in \anc{k_1} \cap \anc{k_2}$ confounding
  $k_1, k_2$.
\end{proposition}
\begin{proof}
  \hl{TODO}
\end{proof}

\begin{corollary}
  If $\gcg$ is strongly causal, then distinct confounders are
  uncorrelated.
\end{corollary}
\begin{proof}
  This follows directly by combining proposition
  \ref{prop:independent_confounders} with proposition \ref{prop:ancestor_properties}.  \hl{I need to update prop} \ref{prop:ancestor_properties} \hl{to note the strong non-correlation property.}
\end{proof}

% \begin{lemma}
%   If $\gcg$ is a strongly causal DAG and $j \in \anc{i}$ then 
% \end{lemma}

@@

  % Should this be written as a lemma?
  \begin{align*}
    x_i(t) &= v_i(t) + \B_{ii}(z) x_i(t) + \sum_{k \in \pa{i}}\B_{ik}(z) x_k(t)\\
           &= \underbrace{v_{a_0}(t) + \B_{a_0a_0}(z) x_i(t) + \sum_{\substack{k \in \pa{a_0} \\ k \ne a_1}}\B_{a_0 k}(z) x_k(t)}_{\defeq \wtalpha{a_0}{a_1}} + \B_{a_0a_1}(z)x_{a_1}(t)\\
           &= \wtalpha{a_0}{a_1} + \B_{a_0a_1}(z)\big[\wtalpha{a_1}{a_2} + \B_{a_1a_2}(z)x_{a_2}(t) \big]\\
           &\overset{(a)}{=} \cdots\\
           &= \sum_{\ell = 0}^r \underbrace{\Big(\prod_{m = 0}^{\ell - 1} \B_{a_m a_{m + 1}}(z) \Big)}_{\defeq F_\ell(z)} \wtalpha{a_\ell}{a_{\ell + 1}} + \Big(\prod_{m = 0}^{r}\B_{a_m a_{m + 1}}(z)\Big)x_{a_{r + 1}}(t)\\
           &= \sum_{\ell = 0}^r F_\ell(z) \wtalpha{a_\ell}{a_{\ell + 1}} + F_{r + 1}(z) x_j(t)
  \end{align*}

@@

  but this is impossible since $x_i(t) \not \in \H_{t - 1}^{(i)}$.
\end{proof}

@@

\subsection{Persistent Systems}
\label{sec:persistent_systems}
In section \ref{sec:strongly_causal_graphs} we obtained a converse to
the part $(a)$ of proposition \ref{prop:ancestor_properties} via the
notion of a strongly causal graph topology.  In this section, we
complete a converse by adding the additional requirement we refer to
as ``persistence''.  This condition requires that each node maintains
some ``memory'' of the past.

\begin{definition}[Lag Function]
  Given a caual filter $\B(z) = \sum_{\tau = 0}^\infty b(\tau)z^{-\tau}$
  define

  \begin{align}
    \tau_0(\B) &= \text{min}\{\tau \in \Z_+\ |\ b(\tau) \ne 0\},\\
    \tau_{\infty}(\B) &= \text{sup}\{\tau \in \Z_+\ |\ b(\tau) \ne 0\}.\\
  \end{align}

  i.e. the ``first'' and ``last'' coefficients of the filter $\B(z)$,
  where $\tau_\infty(\B) = \infty$ if the filter has an infinite
  impulse response.
\end{definition}

\begin{definition}[Persistent]
  We will say that a WSS process $x(t)$ with Granger-causality graph
  $\gcg$ is \textit{persistent} if for every $i \in [n]$, every
  $k \in \anc{i}$ we have $\tau_\infty(\A_{ik}) = \infty$.
\end{definition}

\begin{remark}
  There are a multitude of conditions which would guarantee $x(t)$ is
  persistent, for innstance, any ARMA system with a non-zero
  autoregressive component is likely to be persistent due to recursion
  in the coefficients naturally lead to an $\mathsf{MA}(\infty)$
  representation.

  In particular, any system with a finite impulse response
  (i.e. $\mathsf{VAR}(q)$ models) is likely to be persistent.  These
  models are common in applications of Granger-causality.

  Moreover, persistence is not the weakest condition necessary for the
  results of this section, but we opt to avoid unnecessarily
  complicated definitions.
\end{remark}

\begin{lemma}
  \label{lem:time_lag_cancellation}
  Suppose $v(t)$ is a scalar sequence with unit variance and zero
  autocorrelation and let $\A(z), \B(z)$ be nonzero and strictly
  causal (i.e. $\tau_0(\A) \ge 1$) linear filters.  Then,

  \begin{equation}
    \inner{F(z)\A(z)v(t)}{\B(z)v(t)} = 0\ \forall \text{ strictly causal filters } F(z)
  \end{equation}

  if and only if $\tau_0(\A) \ge \tau_\infty(\B)$.
\end{lemma}
\begin{proof}
  We have

  \begin{align}
    \inner{\A(z)v(t)}{\B(z)v(t)} &= \sum_{\tau = 1}^\infty \sum_{s = 1}^\infty a(\tau)b(s)\E[v(t - s)v(t - \tau)]\\
    &= \sum_{\tau = \text{max}(\tau_0(\A), \tau_0(\B))}^{\text{min}(\tau_\infty(\A), \tau_\infty(\B))} a(\tau) b(\tau)\\
  \end{align}

  due to the uncorrelatedness assumptions on $v(t)$.  This expression
  is $0$ if and only if $\tau_0(\A) \ge 1 + \tau_\infty(\B)$ or if
  $\tau_0(\B) \ge 1 + \tau_\infty(\A)$ or if the coefficients are
  orthogonal along the common support.

  Specializing this fact to $\inner{F(z)\A(z)v(t)}{\B(z)v(t)} = 0$ we
  see that the coefficients cannot be orthogonal for every choice of
  $F$, and that $\text{sup}_F \tau_\infty(F\A) = \infty$, leaving only
  the possibility that

  \begin{align*}
    \tau_0(F\A) \ge 1 + \tau_\infty(\B) \forall F &\overset{(a)}{\iff} \tau_0(\A) \ge 1 + \tau_\infty(\B) - \underset{F}{\text{min }} \tau_0(F)\\
    &\overset{(b)}{\iff} \tau_0(\A) \ge \tau_\infty(\B),
  \end{align*}

  where $(a)$ follows since $\tau_0(F\A) = \tau_0(F) + \tau_0(\A)$,
  and $(b)$ since $\text{min}_F\ \tau_0(F) = 1$.
\end{proof}

\begin{corollary}
  \label{cor:time_lag_cancellation}
  For $k \in \anc{i} \cap \anc{j}$ we have

  \begin{align*}
    \linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} &= 0\ \forall \text{ strictly causal } F(z)\\
    \iff \inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} &= 0\ \forall \text{ strictly causal } F(z)\\
    \iff \tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})
  \end{align*}
\end{corollary}
\begin{proof}
  The final equivalence follows immediately from the Lemma.  For the first equivalence we have

  \begin{align*}
    \linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} &= 0\ \forall \text{ strictly causal } F(z)\\
    \iff \inner{F(z)A_{jk}(z)v_k(t)}{x_i(t - \tau)} &= 0\ \forall \tau \ge 1, \text{ strictly causal } F(z),
  \end{align*}

  which can be expanded by equation \eqref{eqn:ancestor_expansion} to
  obtain (after cancellinng all ancestors of $i$ other than $k$)

  \begin{equation*}
    \inner{F(z)A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t - \tau)} = 0\ \forall \tau \ge 1, \text{ strictly causal } F(z),
  \end{equation*}

  which by the Lemma is equivalent to $\tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})$ as stated.
\end{proof}

\begin{proposition}
  Suppose $\gcg$ is a strongly causal DAG and that $x(t)$ is persistent DAG, then if $k$
  confounds $i, j$ we have $i \pwgc j$ and $j \pwgc i$.
\end{proposition}
\begin{proof}
  We will show that $j \pwgc i$, the other being symmetric.  First
  note also that by proposition \ref{prop:sc_graph_common_anc} we
  cannot have $i \in \anc{j}$ or $j \in \anc{i}$ and therefore every
  $k \in \anc{i}\cap\anc{j}$ will be a confounder.

  It is sufficient to show that $\exists \psi \in \H_{t - 1}^{(j)}$
  such that

  \begin{equation*}
    \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}} \ne 0.
  \end{equation*}

  To this end, let $F(z)$ be an arbitrary but strictly causal linear,
  and equation \eqref{eqn:ancestor_expansion} to $x_i(t)$ and
  $\psi \defeq F(z)x_j(t)$:

  \begin{align*}
    &\inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{x_i(t) - \linE{x_i(t)}{\H_{t - 1}^{(i)}}}\\
    &\overset{(a)}{=} \inner{\psi - \linE{\psi}{\H_{t - 1}^{(i)}}}{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}\\
    &\overset{(b)}{=} \inner{\sum_{k \in \anc{j}}\big(F(z)\A_{jk}(z)v_k(t) - \linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}\big)}{\A_{ii}(z)v_i(t) + \sum_{k \in \anc{i}}\A_{ik}(z)v_k(t)}\\
    &\overset{(c)}{=} \sum_{k \in \anc{i}\cap\anc{j}}\Big(\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} - \inner{\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}}{\A_{ii}v_i(t)}\\
    &- \sum_{\ell \in \anc{i}}\inner{\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}}}{\A_{i\ell}(z)v_\ell(t)}\Big)
  \end{align*}

  where in $(a)$ we have removed the $\linE{x_i(t)}{\H_{t - 1}^{(i)}}$
  term via the orthogonality principle, in $(b)$ there is no
  $F(z)\A_{jj}(z)v_j(t)$ term since due to $j \not\in \anc{i}$ it is
  orthogonal to $\H_t^{(i)}$.  Finally, $(c)$ follows by applying
  orthogonality properties of $v(t)$, as well as the fact that
  $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} = 0$ for
  $k \not \in \anc{i}$.  Note that
  $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} \in \H_{t - 1}^{(i)}$
  and thence there is in general no cancellation in the final term
  above for $\ell \in \anc{i}$.

  Evidently, (\hl{This is clearly not immediately evident}) this is
  $0$ for each $F$ if and only if
  $\linE{F(z)\A_{jk}(z)v_k(t)}{\H_{t - 1}^{(i)}} = 0$ and
  $\inner{F(z)\A_{jk}(z)v_k(t)}{\A_{ik}(z)v_k(t)} = 0$, which by
  Collary \ref{cor:time_lag_cancellation} occurs if and only if
  $\tau_0(\A_{jk}) \ge \tau_\infty(\A_{ik})$, which is impossible
  since in general $\tau_0(\A_{jk}) < \infty$ and in a persistent
  graph $\tau_\infty(\A_{ik}) = \infty$.
\end{proof}

@@

    \BlankLine
    \caption{Pairwise Graph Recovery}
    \label{alg:pwgr}
    % \TitleOfAlgo{Pairwise Graph Recovery}
    \Input{Pairwise Granger-causality relations between $n$ processes whose joint Granger-causality relations are known to form a strongly causal DAG $\gcg$.}
    \Output{Edges $E = \{(i, j) \in [n] \times [n]\ |\ i \gc j \}$ of the graph $\gcg$.}
    \Initialize{$S_0 = [n]$  \texttt{\# unprocessed nodes}\\
      $E_0 = \emptyset$  \texttt{\# edges of }$\gcg$\\
      % $P_0 = \emptyset$  \texttt{\# layer by layer driving nodes}\\
      $k = 0$ \texttt{\# a counter used only for notation}}
    \BlankLine
    $W \leftarrow \{(i, j)\ |\ i \pwgc j, j \npwgc i\}$  \texttt{\# potential edges}\\
    \While{$S_k \ne \emptyset$}{
      $P_k \leftarrow \{i \in S_k\ |\ \forall s \in S_k\ (s, i) \not\in W\}$  \texttt{\# parents of current layer}\\
      % $P_k \leftarrow P_{k - 1} \cup P_k$ \texttt{\# Expand parent set}\\
      $S_{k + 1} \leftarrow S_k \setminus P_k$ \texttt{\# remove from } $S$\\

@@

      \For{$r = 0, \ldots, k$}
      {
        \label{alg:inner_loop}
        $D_{kr} \leftarrow \{(i, j) \in P_{k - r} \times C_k\ |\ (i, j) \in W,\ \text{no } i \rightarrow j \text{ path in } E_k \}$\\%  \texttt{\# edges incident on }$C_k$\\
        $E_k \leftarrow E_k \cup D_{kr}$ \texttt{\# add edges to }$E_k$\\
        % $W_{k + 1} \leftarrow W_k \setminus D_{kr}$  \texttt{\# remove edges from consideration}\\
      }
      % $W_{k + 1} \leftarrow W_k \setminus \big(\bigcup_{r = 0}^k D_{kr}\big)$  \texttt{\# remove edges from consideration}\\

      \;
      $k \leftarrow k + 1$
    }
    $E \leftarrow E_{k + 1}$\\
    \Return{$E$}
  \end{algorithm}
\end{theorem}

@@

  The algorithm proceeds first with the parentless nodes $1, 2$ on the
  initial iteration where the edge $(1, 3)$ is added to $E$.  On the
  next iteration, the edges $(3, 4), (2, 4), (3, 5)$ are added, and
  the false edges $(1, 4), (1, 5)$ is excluded due to the paths
  $1 \rightarrow 3 \rightarrow 4$ and $1 \rightarrow 3 \rightarrow 5$
  in E.  Finally, edge $(4, 6)$ are added, and the false
  $(1, 6), (3, 6), (2, 6)$ edges are similarly excluded due to the
  ordering of the inner loop.

  \begin{figure}
    \centering
    \caption{Example graph for Algorithm \ref{alg:pwgr}}
    \footnotesize{Black arrows indicate true parent-child
      relations.  Red dotted arrows indicate pairwise causality (due to
      ancestor relations), green dash-dotted arrows indicates
      bidirectional pairwise causality (due to the confounding node
      $1$).  Blue groupings indicate each $P_k$ in Algorithm
      \ref{alg:pwgr}.}
    \label{fig:example_fig3}

@@

\begin{lemma}[Loop Invariant]
  \label{lem:loop_invariant}
  $\forall k \ge 0,\ e \in E_k \Rightarrow e \in \mathcal{E}$.
  \hl{Not sure if this is useful / the right approach.}
\end{lemma}
\begin{proof}
  We will proceed by induction.  Firstly, the base case
  $E_0 = \emptyset$ is trivial, so fix some $k \ge 1$ and consider an
  edge $(i, j) \in \big(\bigcup_{r = 0}^k P_{k - r} \big) \times C_k$
  such that $(i, j) \in W$ (if the edge is not in $W$ then it is not
  in $\mathcal{E}$ by Lemma \ref{lem:W_subset_E}).

\begin{definition}[Depth]
  For our present purposes we will define the \textit{depth} $d(j)$ of
  a node $j$ in $\gcg$ to be the length of the \textit{longest} path
  from a node in $P_0$ to $j$, where $d(j) = 0$ if $j \in P_0$.  It is
  apparent (\hl{right?}) that such a path will always exist in a
  strongly causal graph.  For example, in Figure
  \ref{fig:example_fig3} we have $d(3) = 1$ and $d(4) = 2$.
\end{definition}

\begin{lemma}
  $i \in P_k \iff d(i) = k$ and $j \in S_k \iff d(j) \ge k$.
\end{lemma}
\begin{proof}
  We proceed by induction.  The base case $j \in P_0 \iff d(j) = 0$ is
  by definition, and $j \in S_0 \iff d(j) \ge 0$ is trivial since
  $S_0 = [n]$.  So suppose that the lemma is true up to $k - 1$.

  For $i \in P_k$ we have by construction that there is no $j \in S_k$
  such that $(j, i) \in W$ and therefore no $j \in S_k$ such that
  $j \in \pa{i}$.  Therefore, any parent of $i$ can only be in
  $S_{k - r}$ for $r \ge 1$ and by the induction hypothesis which
  implies that $d(i) \le k$.  But if $d(i) < k$ then $i \in P_{k - r}$
  for some $r \ge 1$ (again by the induction hypothesis) thence
  $d(i) = k$ since $P_k \cap P_{k - r} = \emptyset$ by construction.
